RCubedwrk

### 2017-12-13 R3/R3-WP4/RCubed_wrk_hbe.Rmd
updated Rmd-file testing things and added a file with command lists to hack on wrf-files to makethem nice for esd

---
title: "RCubedwrk"
author: "Helene B. Erlandsen"
date: "November 16, 2017"
output: html_document
---

##GCM Files
downloaded from Norstore using python script .... and a modified version of ....py (provided by MP uib) to be compatible with Python version NN (a change in char byte-char handeling needed specification of utf8 format)

@@

Some cdo nco hacking is done to give the files 2-D latitude and longitude variables (cdo does not like wrf).

@@

## Data Reading & Structure cp from am

```{r "setup"}
#require("knitr")
opts_knit$set(root.dir = "~/git/esd_Rmarkdown/R3/R3-WP4/")
opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE) 
```

### Require the following libraries
```{r required R packages,echo=FALSE}
require(esd)
#require(leaflet)
#require(sp)
#require(rgdal)
```

```{r setpath}
#install.packages("~/git/esd",repos=NULL,type="source")
library(esd)
source('~/git/esd/R/retrieve.R')
rcmdatapath <- '~/storah/RCubed/data/RCM'
gcmdatapath <- '~/storah/RCubed/data/GCMraw'
gcmbcdatapath <- '~/storah/RCubed/data/GCMBC'
eraidatapath<-'~/storah/RCubed/data/ERAI'
RCMmonp <- 'wrf_mon_NorESM_BC_2000_2010.nc' 
GCMBCmon <- 'NorESM_BC4R3_esd_2000_2010.nc'
GCMmon <- 'NorESM_nonBC4R3_esd_2000_2010.nc'
ERApl <- 'EraI_pl_4_esd_mdfa_197901to201706.nc' 
ERAsf <- 'EraI_sfc_4_esd_mdfa_197901to201706.nc'

@@

```{r}
gridn <- par(mfrow=c(4, 5))
RCMt2m <- retrieve.rcm(ncfile=RCMmonp, path=rcmdatapath,param='T2MEAN',verbose = TRUE)
# just testing
map(RCMt2m,FUN ="mean")
par(gridn)
```

```{r}
RCMfw <- retrieve.rcm(ncfile=RCMmonp,path=rcmdatapath,param='fw',verbose = TRUE)
gridn <- par(mfrow=c(4, 5))
# just testing
map(RCMfw,FUN ="mean")
par(gridn)
```

```{r loadGMCBC}
GCMBCt2m <- retrieve(ncfile=GCMBCmon, path=gcmbcdatapath,param='tas',verbose = TRUE)
GCMBCt2m<-subset(GCMBCt2m,is=list(c(-20,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
#GCMBCt2 <- attr(psl.nobc,'units') <- 'Pa'
map(GCMBCt2m,FUN ="mean")
```

>>>>

---
title: "RCubedwrk"
author: "Helene B. Erlandsen"
date: "November 16, 2017"
output:
  pdf_document: default
  html_document: default
---

##GCM Files
downloaded from Norstore using python script .... and a modified version of *wps.py* (provided by MP, UiB) to be compatible with Python 3.6.3 (a change in char byte-char handeling needed specification of utf8-format)

@@

For the dynamically downscalled bias corrected NorESM WRF run some cdo nco hacking is done to give the files 2-D latitude and longitude variables and dimensions x,y, and CF-time (cdo does not like wrf). The same is done to the WRF file with the geograhical features (geo_em). The cdo and nco commands are stored in scripts saved for later.

@@

## Data Reading & Structure (copied parts from Abdelkader)

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = "~/git/esd_Rmarkdown/R3/R3-WP4/")
opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE) 
```

### Require the following libraries
Also added a boolean for update esd (provided that a new version is pulled)
```{r required R packages,echo=FALSE}
needed<-c("ncdf4", "zoo")
update_esd <- FALSE
for (package in needed) { 
  if (!require(package,character.only = TRUE, quietly = TRUE)) { 
      install.packages(package);library(package,character.only = TRUE)
    }
}

if ((!require('esd',character.only = TRUE, quietly = TRUE))| (update_esd) ){ 
      install.packages("~/git/esd",repos=NULL,type="source");library('esd',character.only = TRUE)
    }
#require(esd)  
#require(leaflet)
#require(sp)
#require(rgdal)
```

```{r setpath}
#
library(esd)
#source('~/git/esd/R/retrieve.R')
rcmdatapath <- '~/storah/RCubed/data/RCM'
rcmgeopath <-'~/storah/RCubed/data/RCM/WRF_filetypes'
gcmdatapath <- '~/storah/RCubed/data/GCMraw'
gcmbcdatapath <- '~/storah/RCubed/data/GCMBC'
eraidatapath<-'~/storah/RCubed/data/ERAI'
RCMgeo<-'R3geo_emd03run1.nc'
RCMmonp <- 'wrf_mon_NorESM_BC_2000_2010.nc' 
GCMBCmon <- 'NorESM_BC4R3_esd_2000_2010.nc'
GCMmon <- 'NorESM_nonBC4R3_esd_2000_2010.nc'
ERApl <- 'EraI_pl_4_esd_mdfa_197901to201706.nc' 
ERAsf <- 'EraI_sfc_4_esd_mdfa_197901to201706.nc'

@@

```{r, warning=FALSE}
gridn <- par(mfrow=c(4, 5))
RCMt2m <- retrieve.rcm(ncfile=RCMmonp, path=rcmdatapath,param='T2MEAN',verbose = FALSE)
# just testing
attr(RCMt2m,'greenwich')=TRUE
RCM_oro <- retrieve.rcm(ncfile='geo_em_d03ED.nc',path=rcmgeopath,param='HGT_M',verbose = TRUE)

#map.station(RCM_oro,FUN ="mean",cex=0.3,new=F, is=list(lon=c(3.75,15),lat=c(58,64)))

attributes(RCM_oro) <- NULL
attr(RCMt2m,'altitude') <- RCM_oro

RCMt2m <- subset(RCMt2m, is=list(lon=c(3.75,15),lat=c(58,64)))
map.station(RCMt2m,FUN ="mean",cex=0.3,new=F)
par(gridn)

#check later the stripes

```{r,warning=FALSE}
RCMfw <- retrieve.rcm(ncfile=RCMmonp,path=rcmdatapath,param='fw',verbose = FALSE)
RCMfw <- subset(RCMfw, is=list(lon=c(3.75,15),lat=c(58,64)))
gridn <- par(mfrow=c(4, 5))
# just testing
map.station(RCMfw,FUN ="mean")
par(gridn)
```

```{r loadGMCBC,warning=FALSE,fig.height=32}

GCMBCt2m <- retrieve(ncfile=GCMBCmon, path=gcmbcdatapath,param='tas',verbose = FALSE)
attr(GCMBCt2m,'Calendar') <- 'noleap'
GCMBCt2m<-subset(GCMBCt2m,is=list(c(-10,15),c(45,72)),it=c("2000-12-01","2010-12-01"))

map.field(GCMBCt2m,FUN ="mean", new=F)

```{r loadGMCnonBC,warning=FALSE}
#loading non BC data
GCMt2m <- retrieve(ncfile=GCMmon, path=gcmdatapath,param='tas',verbose = FALSE)
attr(GCMt2m,'Calendar') <- 'noleap'
GCMt2m<-subset(GCMt2m,is=list(c(-10,15),c(45,75)),it=c("2000-12-01","2010-12-01"))

#map.field(GCMt2m,FUN ="mean", new=F)

```{r loadERAI,warning=FALSE}

ERAIt2m <- retrieve(ncfile=ERAsf, path=eraidatapath,param='t2m',verbose =FALSE)
ERAIt2m<-subset(ERAIt2m,is=list(c(-10,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
#GCMBCt2 <- attr(psl.nobc,'units') <- 'Pa'
gridn <- par(mfrow=c(4, 5))
map.field(ERAIt2m,FUN ="mean")
par(gridn)
```

```{r corrfield test, warning=FALSE,eval=FALSE}
# the station object : pca
# the field object : eof
# later check out https://github.com/metno/esd_Rmarkdown/blob/master/KSS/dse4KSS.R
# 4 similar examples, there Y is station values
# LTR try DS on EOF RCM predictand on EOF GCM predictor
# bc EOFs contain pattern of variance biases in different GCMs not very important, and can try same model for all benestad or try leave on out as in Walton
# latr try common eof append [Re-analysis or GCM-BC | GCM raw] -->time  to find common eofs
#need to consider that we have a short time-series. maybe need monthly data 12*9, monthly anomalies, or seasonal 4*9

#Regridding and making seasonal instead of monthly to look at EOF
RCMt2m.regrid <- regrid.station(RCMt2m,is=list(lon=seq(4,15,by=0.1),lat=seq(58,64,by=0.25)),verbose=T)
# OBS if loaded demos The following object is masked from ‘package:esd’: unit, that is why unit in regrid.station does not work
RCMt2m.seasons.regrid <- as.4seasons(RCMt2m.regrid)
# Make EOFs of the RCM 
eof.test <- EOF(RCMt2m.seasons.regrid)
plot(eof.test)
# Make EOF of the seasonal anaomaly (monthly?)
eof.test <- EOF(as.anomaly(RCMt2m.seasons.regrid))
plot(eof.test)
plot(EOF(GCMBCt2m))

eof.gcm <- EOF(as.anomaly(GCMBCt2m))
plot(eof.gcm)
plot(eof.gcm,ip=2)
ds <- DS(eof.test,eof.gcm)
plot(ds) #r=0.79
plot(ds,ip=2)
plot(eof.test,ip=2)

#Slightly better result from BC model (which is boundary for dyn-DS)
#eof.gcmnoBC <- EOF(as.anomaly(GCMt2m))
#dsnoBC <- DS(eof.test,eof.gcmnoBC)
#plot(dsnoBC) #r=0.76

#Also trying ERA-I  really bad bc not source of dyn RCM
eof.eraI <- EOF(as.anomaly(ERAIt2m))
#dsEra <- DS(eof.test,eof.eraI)
#plot(dsEra) #r=0.3
#dev.new(); plot.zoo(subset(pca.y,ip=1:4))
#dev.new(); plot.zoo(subset(pca.y,ip=1:4))

RCMt2m.seasons.eof <- EOF(subset(RCMt2m.seasons.regrid,it='djf'),n=3)
plot(RCMt2m.seasons.eof,ip=1)

```{r pca_on_rcm_seasonal_anomalies}
#Making seasonal and looking at PCA

# Walton method was PCA on each of 60 warming patterns (not warming pattern here) with regional means removed
RCMt2m.ses <-as.4seasons(as.monthly(RCMt2m))
RCMt2m.ses.ano <-as.anomaly(RCMt2m.ses)
RCMt2m.ses.anomaly.pca <- PCA(RCMt2m.ses.ano,n=3)
plot.pca(RCMt2m.ses.anomaly.pca, new=F)
plot.pca(RCMt2m.ses.anomaly.pca, new=F, ip=2)
plot.pca(RCMt2m.ses.anomaly.pca, new=F, ip=3)
attr(RCMt2m.ses.anomaly.pca,'greenwich')=TRUE
#RCMtm.clim <- as.climatology(as.monthly(RCMt2m))
```

To do the converison of PCA to EOF we need to load some extra packages not loaded by defualt into *esd*. We call the demo-scripts for *esd* *gridmap* and *pca2eof* to load the packge *LatticeKrig*, *spam*, *grid*, *maps*, *fields* [*grid::unit* will replace the *esd*-function *unit*]

```{r loadingaux, include=FALSE}

#This this is based on a package from NCAR using kringing taking altitude into account, so need wrf-oro
# since oro is lacking now making dummy oro

#pca2eof uses etop5 min is around 8-10 km, lower than original grid of 4 km about does it matter?
reload_auxpckg=F
if (reload_auxpckg==TRUE){
  demo(package='esd');demo(gridmap,ask=FALSE);demo(pca2eof,ask=FALSE)
}

```{r rcmpca2eof_cca_gcm_eof}
RCMt2m.seasons.eof.from.pca <- pca2eof(RCMt2m.ses.anomaly.pca) 
#Compare not same bca regridding and dummy oro
#dev.new()
plot(RCMt2m.seasons.eof.from.pca)
#dev.new()
plot(RCMt2m.ses.anomaly.pca)

#vis(RCMt2m.ses.anomaly.pca)
#Compare
#dev.new()
#plot(RCMt2m.seasons.eof,ip=2)
#dev.new()
#plot(RCMt2m.seasons.pca,ip=2)

eof.gcm <- EOF(as.anomaly(as.4seasons(GCMBCt2m)))
cca <- CCA(RCMt2m.ses.anomaly.pca,eof.gcm)
plot(cca,new=T) #RCMt2m.pca <- PCA(RCMt2m)

```{r test combine EI noresm}
#know from before that ERAI has poor poor temporal covariance as expected with DS norESM
comb <- combine(as.anomaly(as.4seasons(ERAIt2m)),as.anomaly(as.4seasons(GCMBCt2m)))
ceof <- EOF(comb, n=5)
plot(ceof,new=T)
ds <- DS(RCMt2m.seasons.eof.from.pca,ceof)
plot(ds, new=T)
